---
title: "Week 9"
author: "Eddie Perez"
date: '2022-04-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h3>Fetching data from New York Times APIs</h3>

The New York Times web site provides a rich set of APIs, as described [here](http://developer.nytimes.com/docs).

1. Sign up for at least one API key  
2. Choose an API to search and construct an interface in R to read in the JSON data   
3. Transform the data into an R

# Getting Proper Library if it's not install then install it
```{r, warning=FALSE, message=FALSE}
if (!require('stringr')) install.packages('stringr')
if (!require('dplyr')) install.packages('dplyr')
if (!require('jsonlite')) install.packages('jsonlite')
```

# Importing and Searching and Prepping Key

Using the Top Stories API to look for stories in the Technology section of the NY Times.
Setting up my key for importing articles.

```{r}

key <- "&api-key=Yrr0djuV78dAkxirOY3EwIkbXJbrXtDM" # API key


```


## Retrieve

```{r}
# add url and text required for api request
url <- "https://api.nytimes.com/svc/topstories/v2/technology.json"
url <- paste0(url, "?")
# make api request using link provided by NY Times
feed <- fromJSON(paste0(url, key))

```


# Setting up Data Frame with relevant article

```{r}
article <- feed$results

article <- article %>% 
    filter(section == "technology") %>% 
    select(title, abstract, byline, published_date)


```



## Conclusion

```{r}
knitr::kable(article)
```