---
title: "Project 1"
author: "Eddie Perez"
date: '2022-02-20'
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Project #1
In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

```{r libraries}
# setting proper libraries

library(stringr)
library(knitr)

```

## Importing the data.
I used the read csv function to get my data and paste0 to concatenate because of the spaces on the chess file.

```{r importing Data}
# Assign my Chess Data file located on github so that I can use it with the Read table Function
data = "https://raw.githubusercontent.com/Eperez54/Dat-607/main/Project%201/ChessData.txt"
chessData <- read.csv(paste0(data), header = F)

head(chessData)

```
## Cleaning up

Removing unnecessary data the first four rows doesn't really contain information that we need I decided to truncate 
```{r extracting data}
#omitting the first four lines as they do not have valid data but keeping everything else

chessData <- chessData[-c(1:4),]
head(chessData)

```

## Separating Data

I noticed that both rows could be separated and extracted 

```{r}
player <- chessData[seq(1, length(chessData), 3)]
rating <- chessData[seq(2, length(chessData), 3)]

head(player)
head (rating)
```

Right now I will be separating based on information need for the new chessdata.csv file. Here the skills that I learned from last week's homework came into effect and it was very useful in separating and extracting data, based on patterns

```{r}

pairNumber <- as.integer(str_extract(player, "\\d+"))
player_Name <- str_trim(str_extract(player, "(\\w+\\s){2,3}"))
points <- as.numeric(str_extract(player, "\\d+\\.\\d+"))
opponents <- str_extract_all(str_extract_all(player, "\\d+\\|"), "\\d+")
draw <- str_count(player, "\\Q|D  \\E")
lost <- str_count(player, "\\Q|L  \\E")
Won <- str_count(player, "\\Q|W  \\E")

state <- str_extract(rating, "\\w+")
player_Rating <- as.integer(str_extract(str_extract(rating, "[^\\d]\\d{3,4}[^\\d]"), "\\d+"))

```
## Calculating average 


```{r}
opp_Rating <- length(player)
for (i in 1:length(player))  
  opp_Rating[i] <- round( mean ( player_Rating [as.numeric (unlist( opponents[ pairNumber[i]]))]), digits = 0) 

```

## Creating a new dataframe to hold my final chess data ready for export
```{r}
finalChessData <- data.frame(pairNumber, player_Name, state, points, player_Rating, opp_Rating, Won, lost, draw)
head (finalChessData)
```
## Exporting to a csv file
I use the write to csv file to export my chess data to file chessData.csv

```{r}
write.csv(finalChessData,file = "chessData.csv")
```

## Conclusion

This project was a bit tricky because I knew where I wanted to end up but getting there was hard. Thankfully I used some of string manipulation that we learned from last week lab which helped me get there. I wonder if it is possible to solve this without using string manipulation (Patterns)