---
title: "HW#3"
author: "Eddie Perez"
date: "2/19/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE}
library(tidyverse)
library(stringr)
```

### #1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

### Loading the Data
```{r Exercise #1}
file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
major_list <- read.csv(file, TRUE, ",")
head (major_list)
```
## Searching for Data and Statistics Majors 
searching and storing majors from Data and Statistics

```{r Majors}
Data_Majors <- str_subset(major_list$Major,"DATA")
Statistic_Majors <- str_subset(major_list$Major,"STATISTICS")

print("Below you will find the Data major\n")
Data_Majors

print("\nBelow you will find the Statistic major\n")
Statistic_Majors

```


### #2 Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

```{r Exercise #2}
fruits <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

fruits <- str_extract_all(fruits,'[a-z]+\\s[a-z]+|[a-z]+')
unlist(fruits)

```


### #3 Describe, in words, what these expressions will match:
(.)\1\1 **This will look for the first Character that doesn't start on a new line and see if it repeats twice thereafter**

"(.)(.)\\2\\1" **It will look at the first two letters of a word and see if something matches the inverse**

(..)\1 **This would match character grouping for example a set of words that repeat, for example church the ch would be a good example**

"(.).\\1.\\1" **a specific letter repeated once then another letter repeated once**

"(.)(.)(.).*\\3\\2\\1" **This would match any 3 characters pair repeated 3times**



### #4 Construct regular expressions to match words that:
Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r Exercise #4}
# Setting up my data for following Exercise

x <- c("eleven","banana","Mississippi","hello","today","yesterday","mom","dad")


```


#### Start and end with the same character.

```{r Start and end}
str_view(x, "^(.)(.*)\\1$")
```


#### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r repeated pair}

str_view(x, "([A-Za-z][A-Za-z]).*\\1")

```

#### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r one repeated letter}

str_view(x, "([A-Za-z]).*\\1.*\\1")

```

