---
title: "607 Homework 1"
author: "Eddie Perez"
date: "February 5, 2022"
output: html_document
---

# How much money does NYPD spend on its officers.
As a city worker myself I always hear there is not enough money in the budget. Therfore I decided where the city of NY could save money and spend it on more important angency or public, like streets or even on homeless people.


# Load Some libraries

```{r}
library (devtools)
library (tidyverse)
library(knitr)
```

# Import File from Github 

```{r}
file <- ("https://raw.githubusercontent.com/Eperez54/Dat-607/main/week%201/The%20Marshall%20Project%20-%20NYPD%20Claims%20Settled%20From%201.1.2010%20to%2012.31.2019.csv")


data <- read.csv(file)
data <- data.frame(data) %>% na.omit() # not counting N/A Values
head (data)
```

# Get column headers 

```{r}
colnames(data)
```
## Columns name are understandble no need to rename them.
I want to only see vehicle accidents where NYPD was involved
```{r}
VehicleAccident <- subset(data, data$Claim.Type == "AUTOMOBILE ACCIDENT")
head (VehicleAccident)
```
# Bad Drivers?
How much does NYPD cost the city with bad officer drivers
first we have to change the dollar amount to numeric

```{r}
VehicleAccident$Settlement.Amount <- as.double(gsub('[$,]', '', VehicleAccident$Settlement.Amount))

meancost = mean(VehicleAccident$Settlement.Amount)
occurances = nrow(VehicleAccident)

sprintf("The mean value of settlement is $%.2f", meancost)
sprintf("The total amount of accidents is %.0f", occurances)
sprintf("The highest dollar amount in Settlement is $%.2f", max(VehicleAccident$Settlement.Amount))
```


```{r}
options(scipen=999) #  telling R to avoid all scientific notation by setting options
#(data$Claim.Type)

barplot(table(VehicleAccident$Occurrence.Borough), ylim=c(0,2000))

```

```{r}
barplot(table(VehicleAccident$Occurrence.Borough),                     # Draw barplot with properly aligned legend
        col = 1:nrow(table(VehicleAccident$Occurrence.Borough)),
        legend.text = TRUE, 
        args.legend = list(x = "topright",
                           inset = c(- 0.15, 0)))

```


# Conclusion

I believe that the NYPD could trim down this numbers. In a job where you spend more than 8 hours on the road the vehicle becomes your office. Furthermore I also believe that money spent in settlement could be better used in either training officers better driving skills to avoid accidents or invest in a program where officers who are repeat offenders would be mandated to go into training. If not then use on a different NYC agency.



